Interval Probability and Measurement for Fuzzy Quantum Theories

### 2016-09-27 prob.tex
From: Sabry, Amr A. Sent: Tuesday, September 27, 2016 9:25 To: Tai, Yu-Tsung <yuttai@indiana.edu>; Ortiz, Gerardo <ortizg@indiana.edu>; Hanson, Andrew J. <hansona@indiana.edu>; Liu, Weihua <liuweih@indiana.edu> Subject: Shapley 1965
Hi,

A central piece of information is the theorem of Shapley 1965 cited in
our draft. A few thoughts on that:

* Administrative: Yu-Tsung, you cite [28,29] for the theorem but only
[28] is by Shapley. Also that paper is an internal memorandum but there
is a published version
http://link.springer.com/article/10.1007/BF01753431
that should be cited instead.

* Yu-Ts

\begin{thm}[Shapley~\cite{Shapley1965,GilboaSchmeidler1994}]\label{thm:Shapley}
Every convex interval-valued probability measure has a nonempty core.\qed\end{thm}

@@

\begin{thm}[Bondareva-Shapley~\cite{Shapley1967,Bondareva1968,Grabisch2016}]\label{thm:Bondareva-Shapley}
Let $l$ be the left-end of an interval-valued probability measure~$\bar{\mu}$,
$\bar{\mu}$ has a non-empty core if and only if for any minimal balanced
sets~$\left\{ E_{i}\right\} _{i=1}^{N}$, we have $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\gamma_{i}l\left(E_{i}\right)\le1$,
where $\left\{ E_{i}\right\} _{i=1}^{N}$ is called minimal balanced
sets if they are distinct, $\emptyset\ne E_{i}\subsetneq\Omega$,
and there exist a unique positive coefficient~$\left\{ \gamma_{i}\right\} _{i=1}^{N}$
such that $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1\textrm{ and }\omega\in E_{i}}^{N}}\gamma_{i}=1$
for all $\omega\in\Omega$.\qed\end{thm}

The Bondareva-Shapley theorem can be used to algorithmically decide
non-empty core because the minimal balanced sets can be generated
by an algorithm~\cite{Peleg1965}. However, the number of minimal
balanced sets grows very rapidly with $\left|\Omega\right|$, the
size of $\Omega$~\cite{Shapley1967,Zhao2008}\footnote{\yutsung{Prove the number of minimal balanced sets grow exponentially
or more?}}, so that we have only verified every interval-valued probability
measure has a non-empty core when $\left|\Omega\right|\le5$ so far.\footnote{\yutsung{The program for $\left|\Omega\right|=6$ blows up the memory
again, and the symmetric answer doesn't really work. Hence, currently
only back-track randomized program still searches IVPM with a extremely
slow speed...}} Based on the previous fact, we have the following conjecture.

@@

\begin{example}{[}Three-dimensional quantum 4-value interval-valued
probability measure{]}\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify whether this example is convex or not.}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined as follow, where the state vectors corresponding to 1-dimensional
projector are also illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{enumerate}
\item Let 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps}) & = & \necess\textrm{ ,}
\end{eqnarray*}
 where $\ket{0}$ and $\ket{\ps}$ is illustrated as the red and green
dotted vector, respectively.
\item Consider the red and green circles in figure~\ref{fig:three-dimensional-4-value}.
They are the states orthogonal to $\ket{0}$ and $\ket{\ps}$, respectively,
and can be parametrized as $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. The dotted
half of those states need special treatment, i.e., when where $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define 
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely\textrm{ .}
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi})=\likely$. 
\end{enumerate}
That $\bar{\mu}$ has an empty core can be proved by contradiction.
Assume there is a real-valued probability measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$
for all $P\in\events$. Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]

>>>>

\begin{thm}[Shapley~\cite{Shapley1971}]\label{thm:Shapley}
Every convex interval-valued probability measure has a nonempty core.\qed\end{thm}

@@

\begin{thm}[Bondareva-Shapley~\cite{Shapley1967,Bondareva1968}]\label{thm:Bondareva-Shapley}
Let $l$ be the left-end of an interval-valued probability measure~$\bar{\mu}$,
$\bar{\mu}$ has a non-empty core if and only if for any minimal balanced
sets~$\left\{ E_{i}\right\} _{i=1}^{N}$, we have $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1}^{N}}\gamma_{i}l\left(E_{i}\right)\le1$,
where $\left\{ E_{i}\right\} _{i=1}^{N}$ is called minimal balanced
sets if they are distinct, $\emptyset\ne E_{i}\subsetneq\Omega$,
and there exist a unique positive coefficient~$\left\{ \gamma_{i}\right\} _{i=1}^{N}$
such that $\tensor*[_{\mathbb{R}}]{\sum}{_{i=1\textrm{ and }\omega\in E_{i}}^{N}}\gamma_{i}=1$
for all $\omega\in\Omega$.\qed\end{thm}

The Bondareva-Shapley theorem can be used to algorithmically decide
non-empty core because the minimal balanced sets can be generated
by an algorithm~\cite{Peleg1965,Grabisch2016}. However, the number
of minimal balanced sets grows very rapidly with $\left|\Omega\right|$,
the size of $\Omega$~\cite{Shapley1967,Zhao2008}\footnote{\yutsung{Prove the number of minimal balanced sets grow exponentially
or more?}}, so that we have only verified every interval-valued probability
measure has a non-empty core when $\left|\Omega\right|\le5$ so far.\footnote{\yutsung{The program for $\left|\Omega\right|=6$ blows up the memory
again, and the symmetric answer doesn't really work. Hence, currently
only back-track randomized program still searches IVPM with a extremely
slow speed...}} Based on the previous fact, we have the following conjecture.

@@

\begin{example}[Three-dimensional quantum 4-value interval-valued
probability measure]\label{ex:three-dimensional-4-value} \footnote{\yutsung{Need to verify whether this example is convex or not.}}Given
a three dimensional Hilbert space with an orthonormal basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $.
Consider\emph{ }$\mathscr{I}=\left\{ \imposs,\unlikely,\likely,\necess\right\} $,
$\ket{\ps}=\frac{1}{\sqrt{2}}(\ket{0}+\ket{1})$, $\ket{\ms}=\frac{1}{\sqrt{2}}(\ket{0}-\ket{1})$,
and a quantum $\mathscr{I}$-interval-valued probability measure~$\bar{\mu}:\events\rightarrow\mathscr{I}$
defined as follow, where the state vectors corresponding to 1-dimensional
projector are also illustrated in figure~\ref{fig:three-dimensional-4-value}.
\begin{enumerate}
\item Let 
\begin{eqnarray*}
\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps}) & = & \imposs\\
\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0})=\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps}) & = & \necess\textrm{ ,}
\end{eqnarray*}
 where $\ket{0}$ and $\ket{\ps}$ is illustrated as the red and green
dotted vector, respectively.
\item Consider the red and green circles in figure~\ref{fig:three-dimensional-4-value}.
They are the states orthogonal to $\ket{0}$ and $\ket{\ps}$, respectively,
and can be parametrized as $\ket{0_{\theta,\gamma}^{\perp}}=\rme^{\rmi\gamma}\sin\theta\ket{1}+\cos\theta\ket{2}$
and $\ket{\ps_{\theta,\gamma}^{\perp}}=-\rme^{\rmi\gamma}\sin\theta\ket{\ms}+\cos\theta\ket{2}$,
where $0\le\theta\le\frac{\pi}{2}$ and $0\le\gamma<2\pi$. The dotted
half of those states need special treatment, i.e., when where $0\le\theta<\frac{\pi}{2}$
and $0\le\gamma<\pi$, we define 
\begin{eqnarray*}
\bar{\mu}\left(\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \likely\\
\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{0_{\theta,\gamma}^{\perp}}\right)=\bar{\mu}\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\ps_{\theta,\gamma}^{\perp}}\right) & = & \unlikely\textrm{ .}
\end{eqnarray*}
\item Otherwise, $\bar{\mu}(\proj{\psi})=\unlikely$ and $\bar{\mu}(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}\proj{\psi})=\likely$. 
\end{enumerate}
That $\bar{\mu}$ has an empty core can be proved by contradiction.
Assume there is a real-valued probability measure satisfying $\mu_{\rho}^{B}(P)\in\bar{\mu}(P)$
for all $P\in\events$. Because $\mu_{\rho}^{B}(\proj{0})\in\bar{\mu}(\proj{0})=\imposs$
and $\mu_{\rho}^{B}(\proj{\ps})\in\bar{\mu}(\proj{\ps})=\imposs$,
we must have $\mu_{\rho}^{B}(\proj{0})=\mu_{\rho}^{B}(\proj{\ps})=0$
so that $\mu_{\rho}^{B}=\mu_{\ket{2}}^{B}$. However, 
\[
\mu_{\ket{2}}^{B}(\proj{2})=1\notin\unlikely=\bar{\mu}(\proj{2})\textrm{ .}
\]
